%!TEX root=index.tex
\section{Conclusion}
The general insurance industry is under financial risk due to losses incurred from the frequency and severity of catastrophic weather events. Unless the insurers can effectively balance high-risk policies with lower risk ones, their business model may not be sustainable in the future. ClimatEdge\index{ClimatEdge} has a number of planned offerings around severe weather probabilities that will give insurance underwriters the additional data necessary to categorize the risk level associated with policy offerings.\\

Initial offerings can be developed around tornado and hail probabilities without investing in any computational resources other than a modern server with several terabytes of storage. From an implementation point of view, a hail or tornado analytic would have enough similar computational requirements and data storage capacity\footnote{The creation of a proprietary hail index, as opposed to buying or partnering, would require more volume, greater variety, and faster velocity.} that considering them separately did not add qualitative substance. Hail forensics would be even simpler, just displaying the presence or absence of hail for any recent time frame in a specified location. However as research moves forward into using the \gls{merra} data to reduce uncertainty across offerings, the volume increases substantially. Even with more complex visualizations such as maps or real-time alerting for forensics, it is unlikely that these offerings will ever meet the velocity, volume, or variety necessary of a big data solution. Both global and domestic flood analytics will require a substantially more complex framework in order to succeed. Flood will have petabytes of satellite, radar, and news data along with the complex algorithms necessary to distill that data into usable probabilities. Unless a scalable and elastic framework is put into place, more complex offerings within ClimatEdge\index{ClimatEdge} will have difficulty succeeding.\\

A big data\index{big data} framework that is both flexible and loosely coupled has the greatest chance of meeting future offering requirements while allowing various technology components to be upgraded. Many commercially available offerings are dependent on particular technologies that have well-established strengths and weaknesses. If the framework presentation layer is dependent on a particular storage layer implementation, those two layers are forever tightly coupled and unable to be separated without a major overhaul. While it is expected that a certain amount of coupling is needed\footnote{It is possible to develop abstract interfaces between the layers which minimize the impact of switching technologies.}, e.g. the presentation layer may need to know that certain data structures exist within the data store, it should be possible to implement those structures in Cassandra\index{Cassandra} or MySQL\index{MySQL} as determined by the solution architects.\\

There are several areas of future research centering on how best to store publicly available data sets. There is a dizzying array of tradeoffs in deciding whether to implement remote \gls{api}s, clone the data sets internally, or develop and deploy big data\index{big data} appliances at partner agencies. What is even more daunting is that the solutions will likely vary based on the data set and the planned usage of the resulting analysis, leading to the problem of what works well for climate data, may not work well for healthcare or other commercial big data services.\\

\textbf{DISCLAIMER}\\

The information, views and opinions expressed in this paper constitute solely the authors’ views and opinions and do not represent in any way CSC’s official corporate views and opinions. The authors have made every attempt to ensure that the information contained in this paper has been obtained from reliable sources. CSC is not responsible for any errors or omissions or for the results obtained from the use of this information. All information in this paper is provided “as is,” with no guarantee by CSC of completeness, accuracy, timeliness or the results obtained from the use of this information, and without warranty of any kind, express or implied, including but not limited to warranties of performance, merchantability and fitness for a particular purpose. In no event will CSC, its related partnerships or corporations, or the partners, agents or employees thereof be liable to you or anyone else for any decision made or action taken in reliance on the information in this paper or for any consequential, special or similar damages, even if advised of the possibility of such damages.\\

Copyright \copyright~2014 Computer Sciences Corporation. All rights reserved.
